Brownian dynamics simulation of cytoplasmic dynein's powerstroke reproduces bidirectional stepping

### 2019-01-17 papers/paper/paper.tex
Make some additions to paper text, and replace the onebound/bothbound formalism with prestroke and poststroke to fit better with other papers

%% Several computational models have been created to explain dynein's motility. \textit{Imamula et. al.} define a chemical transition model which defines all the MT- and nucleotide-bound states dynein goes through as it walks \cite{imamulamodel}. \textit{Sarlah et. al} simulate a model obeying rate constants from \textit{Imamula}, and replicate dynein's stepping trajectory \cite{sarlahmodel}. \textit{Zheng} uses normal mode analysis to simulate the dynein motor's transition from pre to post-powerstroke \cite{normalmodes}. To our knowledge, no computational model exists which takes into account the microscopic dynamics of protein-water interactions to test whether diffusion is enough for dynein motility. Here we show how a simple Brownian dynamics model can be used to test the mechanochemical cycle picture of dynein processivity.\\

@@

This mechanochemical cycle of chemical transitions is well supported by evidence \cite{citemore}. However, there is still relatively little known about the dynamic mechanism of domain motion dynein uses to achieve its forward motion by transitioning between these states \cite{lippert}. There exist models theorizing how dynein's domains move during its forward step, such as the winch \cite{} and powerstroke \cite{} models, which can reproduce experimental data \cite{}. However, to our knowledge, no studies exist which study if either of these dynamic mechanisms is feasible for the dynein motor to produce. Existing computational models of walking use chemical rate transitions and assume independence of steps, without simulating the precise protein dynamics \cite{}. Molecular dynamics simulations exist, but are of too short a timescale to describe dynein's forward stroke \cite{}. It is necessary to model dynein dynamically at a large timescale to fully understand how feasible models like the winch and powerstroke are. Further, such long-timescale dynamic models could be used to find what elements of dynein are sufficient to produce forward motion. Further, an accurate dynamical model would allow us to test if certain dynamical elements, such as electrostatic interactions with the microtubule \cite{longrangemt}, ring stacking \cite{sarlahmodel}, discrete microtubule binding sites \cite{trottmodel}, or elasticity of the stalk \cite{burgess-paper} are necessary for dynein's walk.

@@

\begin{figure}[tbhp]
\centering
\includegraphics[width=\linewidth]{figures/model-cartoon-simple.pdf}
\caption{\textbf{Basic geometric model in pre-powerstroke and post-powerstroke states.} \textit{Left: } Pre-powerstroke model of dynein with one MTBD bound to microtubule and the other free. Angles are measured from the $+\hat{x}$ direction. \textit{Right: } Post-powerstroke model with both MTBDs bound to microtubule. MTBD angles are relative to $+\hat{x}$ direction, and motor angles are between stalk and tail. Lengths and radii not to scale.}
\label{fig:model}
\end{figure}

\begin{figure}[tbhp]
\includegraphics[width=0.4\linewidth]{../../plots/burgess-model-figure.pdf}%
\includegraphics[width=0.6\linewidth]{../../plots/grotjahn-model-figure.pdf}%
%% \includegraphics[width=0.3\linewidth]{../../plots/crystal-model-figure.pdf}%
\caption{\textbf{Scale model compared with experimental images.} \textit{Left:} Model with angles at pre-powerstroke equilibrium superimposed over axonemal dynein c heavy chain micrograph in the pre-powerstroke ADP-Vi-bound state \cite{burgess-paper}. Interdomain lengths and radii scaled to micrograph scale bar. No microtubule is present, so $\theta_{nb}$ is arbitrary. \textit{Right: } Model with angles at pre-powerstroke equilibrium superimposed over density map of full-length native cytoplasmic dynein with accessory tail proteins (map ID EMD-7000) \cite{grotjahn}. Interdomain lengths and radii scaled to density map.}
\label{fig:micrographs}
\end{figure}

@@

  \begin{minipage}{\linewidth}
    \small
    \textbf{Table 1. Parameters used in simulation.} $c_b$, $c_m$ and $c_t$ are the MTBD, motor and tail spring constants, respectively. $k_b$ and $k_{ub}$ are the rates of pre-powerstroke to post-powerstroke and post-powerstroke to pre-powerstroke transitions, respectively. $L_s$ and $L_t$ are the lengths of the stalk and tail interdomain linkers. $c$ is the tension-gating factor. The $\theta$ values are the equilibrium values for each angle, where $\theta_m^{\mbox{pre}}$ is the equilibrium angle for the unbound motor in pre-powerstroke, whereas $\theta_m^{\mbox{post}}$ is the equilibrium angle for the bound pre-powerstroke motor and both post-powerstroke motors. $R_b$, $R_m$ and $R_t$ are the MTBD, motor and tail radii. Parameters used for all simulations unless otherwise noted.
  \end{minipage}
  \label{tab:params}

@@

\section{Model}
This dynein model is meant to capture the coarse-grained structural, dynamical and chemical properties of the full dynein complex while remaining mathematically simple. The structural features of dynein are captured through a two-dimensional geometric model of circular domains connected by rigid rods, shown in Figure \ref{fig:model}. The dynamics of dynein are captured by imposing both equilibrium and Brownian forces on each circular domain of the model. The chemical properties of dynein are captured by sporadically transitioning the model between two states: a ``bothbound'' state where both binding domains are fixed to the microtubule, and a ``onebound'' state where one foot is free to diffuse, as has been done before \cite{zhaomodel}. The bothbound state is meant to represent the post-powerstroke pre-ADP release state, and the onebound state represents the post pre-stroke pre-MT binding state, during which ATP is hydrolyzed and phosphate is rejected. These two states were chosen because ADP release is thought to be the rate limiting step for the whole cycle \cite{holzbaur1989}, and the recovery stroke is expected to happen soon after MT unbinding \cite{mogamirate}, meaning the post-recovery conformation should dominate when unbound from MT. We constrain our model to move through only the most likely mechanochemical cycle according to \textit{Tsygankov et al} \cite{tsygankovscheme}, where a head remains in the ADP-bound post-powerstroke state while the other head goers through the entire mechanochemical cycle, neglecting other possible cycles. Each state feels equilibrium forces on its domains meant to mimic the structural changes which occur during the recovery stroke and powerstroke \cite{burgess-paper, burgessknight}. In particular, the equilibrium angles controlling a motor's tail-motor-MTBD angle change to kick the stalk-MTBD further from the tail.

@@

Values for $R$ and $L$ were acquired by aligning the dynein model with EM micrographs of the full dynein complex \cite{burgess-paper,grotjahn}, as shown in Figure \ref{fig:micrographs}. See Table \ref{table:params} for these values.

\subsection*{Equilibrium forces}
The onebound model has four degrees of freedom: $\theta_{bb}, \theta_{bm}, \theta_{um}$ and $\theta_{ub}$. The bothbound model has two: $\theta_{nm}$ and $\theta_{fm}$. The model is free to vary across all of these degrees of freedom, barring steric interactions with the microtubule. These angles describe a planar model which is not capable of transverse motion. To capture the gross dynamics and interdomain oscillations of the protein, restoring forces are imposed on each angle about equilibrium angles $\theta^{eq}_{bb}, \theta^{eq}_{bm}, \theta^{eq}_{um}, \theta^{eq}_{ub}$ for onebound and $\theta^{eq}_{nm}$ and $\theta^{eq}_{fm}$ for bothbound. Each domain $i$ has a harmonic energy $c_i\left(\theta_i-\theta^{eq}_i\right)^2$ which exerts a restoring force pushing it back to equilibrium, where $c_b$, $c_m$ and $c_t$ are spring constants for each domain. These equilibrium angles are taken to be the average angle occupied by the protein in solution, and are thus extracted from EM micrographs, shown in Figure \ref{fig:micrographs}. These angles were extracted from axonemal dynein, but structural similarity suggests they should be applicable to cytoplasmic dynein as well \cite{dynein-c-paper}. Electron microscopy studies using ADP-Vi to freeze dynein in the pre-powerstroke state allow for both pre- and post-stroke equilibrium angles to be found \cite{burgess-paper}.

@@

\subsection*{Binding}
Conformational changes on nucleotide binding, also known as the mechanochemical cycle \cite{cianfrocco}, are captured by transitioning the model between the onebound and bothbound states. Binding can occur when the onebound model's unbound foot diffuses near enough the microtubule. When within a distance of $1 A$, binding occurs at a rate of $k_b$ (Table \ref{table:params}). On binding, the onebound model changes state to a bothbound model while retaining the positions of its domains. The previously unbound domain is transported to the microtubule.

\subsection*{Unbinding}
Unbinding occurs when a bothbound MTBD unbinds from the microtubule. This event can occur for either MTBD, and each domain has its own unbinding rate $k_{ub}$. To account for the tension-gating proposed by Yildiz \textit{et al} \cite{yildiz, yildizcleary} to allow for interhead communication and bias the motor forward, the parameter $c$ is used to alter the unbinding rates of either MTBD. For domain $i \in \{n, f\}$ with binding domain angle $\theta_{ib}\left(\theta_{nm}, \theta_{fm}\right)$, unbinding rate is given by $k_{ub} = k^0_{ub}e^{-c\left(\theta_{ib}-\theta^{eq}_{ib}\right)}$. This will bias the model to unbind more preferentially for the lagging MTBD due to a higher likelihood for $\theta^{eq}_{ib} > \theta_{ib}$ for the lagging head. Values used for parameters $k^0_{ub}$ and $c$, along with $k_b$, are shown in Table \ref{table:params}.

The ``prestroke'' and ``powerstroke'' events in the mechanochemical cycle \cite{cianfrocco} are captured by different motor equilibrium angles between the onebound and bothbound states. The onebound ``prestroke'' state has an unbound motor angle closer to $\pi$ than the bothbound motor\cite{burgess-paper}; thus, on unbinding, the unbound MTBD should experience a restoring force effecting the ``prestroke'' conformational change. On rebinding, the motors transition back to the poststroke equilibrium, producing restoring forces in the direction of the ``powerstroke.''

@@

  \section*{Fitting $k_b$ and $k_{ub}$}
  To fit the binding and unbinding rate constants $k^0_b$ and $k^0_{ub}$ a simple theoretical model for stepping was used, similar to one previously published \cite{myosindutyratio}. First, it is assumed that the unbinding rate of an MTBD does not depend on the binding state of the other MTBD, that is, that there is no inter-head coordination. Then the dissociation rate of the whole complex from the microtubule, $k_{dis}$, is estimated as $k_{dis} = P(ob)*k_{ub} = \frac{\bar{t}_{ob}}{\bar{t}_{step}} * \frac{1}{\bar{t}_{bb}}$, where $P(ob)$ is the probability of a dynein complex being in the onebound state, $\bar{t}_{ob}$ and $\bar{t}_{bb}$ are the expected durations for a onebound and bothbound state, and $\bar{t}_{step} = \bar{t}_{ob} + \bar{t}_{bb}$ is the expected step duration. Plugging in the last definition yields $k_{dis} = \frac{\bar{t}_{step} - \bar{t}_{bb}}{\bar{t}_{step}\bar{t}_{bb}}$. Introducing a new variable $\bar{t}_{run}$, the average run length of in vivo dynein, and using the relation $\bar{t}_{run} = k_{dis}^{-1}$, a new relation is found for bothbound time: $\bar{t}_{bb} = \frac{\bar{t}_{run}\bar{t}_{step}}{\bar{t}_{run}+\bar{t}_{step}}$. Using the definition of $\bar{t}_{step}$ a similar relation is found for onebound time: $\bar{t}_{ob} = \frac{\bar{t}_{step}^2}{\bar{t}_{step}+\bar{t}_{run}}$. Finally, using the relation $\bar{t}_{step} = \frac{v}{L_{step}}$ for dynein velocity $v$ and expected tail step length $\bar{L}_{step}$, the following equations are found for onebound and bothbound time in terms of experimental observables of velocity, run time and step length:

@@

  For estimating onebound and bothbound times from Yildiz \textit{et al}\cite{yildiz}, a velocity of $124 nm/s$ is used. A study with a similar velocity is used to estimate $\bar{t}_{run} = 1060 nm / 134 nm / s = 7.9s$ \cite{weihongpaper}. These observables, along with $\bar{L} = 8nm$, yield estimations $\bar{t}_{ob} = 5.2*10^{-4}s$ and $\bar{t}_{bb} = 6.4 * 10^{-2}s$.

  \section*{Onebound motion equations derivation}
  Brownian Dynamics was used to describe the motion of each domain, indexing with $i \in \{bb, bm, t, um, ub\}$, corresponding to bound-binding, bound-motor, tail, unbound-motor, unbound-binding:

@@

  \section*{Bothbound motion equations derivation}
  Brownian dynamics was also used for the bothbound case. Due to the added constraint of both binding domains being fixed to the microtubule, the model is fully described by angles $\theta_{nm}$ and $\theta_{fm}$; see Figure \ref{fig:model}. The positions of non-bound domains are defined by the following equation for $i \in \{nb, nm, t, fm, fb\}$, corresponding to near-binding, near-motor, tail, far-motor and far-binding domains:

@@

  \section*{Simulating the model}
  Simulations began in the bothbound equilibrium state. The model was time-evolved via Euler's Method using analytic expressions for $\{\dot{\theta}_{bb}, \dot{\theta}_{bm}, \dot{\theta}_{um}, \dot{\theta}_{ub}\}$, using a timestep of $dt = 10^{-10}s$. In the onebound state, during each timestep the model had a $k_b*dt$ probability of transitioning into the bothbound state, where $k_b = k_b^0\left(1-H\left(X_{uby}-0.1\right)\right)$, where H is the Heaviside function. On transition to bothbound, proper $\theta_{nm}$ and $\theta_{fm}$ angles and proper interhead separation $L$ were calculated to yield a bothbound model with identical cartesian positions to the old bothbound, except with the previously unbound binding domain moved to $y=0$. Once in bothbound, the model was time-evolved via Euler's Method using analytic expressions for $\{\dot{\theta}_{nm}$ and $\dot{\theta}_{fm}\}$ and a timestep of $dt=10^{-10}s$. Each timestep, the bothbound model had a $k_{ub}*dt$ chance of transitioning to onebound, where $k_{ub} = k^0_{ub}e^{-c(\theta-\theta_{eq})}$. On transitioning, proper values of $\{\theta_{bb}, \theta_{bm}, \theta_{um}, \theta_{ub}\}$ were calculated to create a onebound model with identical cartesian positions to the old bothbound model.

@@

Motion equations were derived for the pre-powerstroke and post-powerstroke states of the model. The model was simulated by starting in the equilibrium post-powerstroke position and using Euler's method to time-evolve the model, using a timestep of 1e-10 \textit{s}. The model was transitioned between post-powerstroke and pre-powerstroke states using unbinding and binding rates. Model parameters were estimated from various experimental results. Spring constants were chosen to fit the model stepping size to single-molecule head-labelled experiments\cite{yildizpaper}, as outlined in Figure S\ref{fig:supp-model-fit}. Model rate constants were fit to motor velocities from the same experiment, along with theoretical calculations for the prestroke and poststroke times outlined in the Methods section. The tension-gating parameter $c$ was fit to the leading/lagging unbinding likelihood versus interhead separation from the same experiment, as shown in Figure \ref{fig:tensiongating}. Thirty-two 5-second simulations were run, each initialized with a different random seed. Each simulation was forward directed, took both forward and backward steps, and displayed stepping dynamics similar to that seen in experiment.

@@

The model was then compared to TIRF studies of dynein stepping in DeWitt \textit{et. al.} \cite{yildizpaper}. Figure \ref{fig:behavior}.a shows 8 representative model trajectories compared to experimental dynein. The eight simulations track experimental very well over 5 seconds, and average to the same position. The eight simulations end up distributed over a 200 \textit{nm} area, showing the variation in the model stepping extends to a variation in velocity in the long term. The step size distribution for the 32 simulations are shown in Figure \ref{fig:behavior}.b). The model is capable of taking both positive and negative steps, with steps extending from -40 \textit{nm} to 50 \textit{nm}, matching experimental dynein. The most probable step is 12 \textit{nm}, with an average step size of 6.2 \textit{nm}. This matches the peak experimental step size window of 7-13 \textit{nm}. The pre-powerstroke and post-powerstroke state occupation times are shown in Figure \ref{fig:behavior}.c and d.). The model prefers long pre-powerstroke times, but is capable of taking very fast steps. The post-powerstroke time has a heavy-tailed distribution. Experimental technique does not resolve time spent in pre- and post-powerstroke, but we outline a technique for calculating these values for experimental trajectories in Methods. These theoretical values correspond well to model behavior, and square with experimental findings that dynein is almost always in the post-stroke state \cite{imaiburgess}.

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\linewidth]{../../plots/paper_model_behavior}
\caption{\textbf{Stepping dynamics of model.} \textit{Top: } Average position of near and far MTBDs over 5 seconds of simulation and 8 random seeds. Trajectory compared with single-motor-labelled cytoplasmic dynein from DeWitt \textit{et. al.} \cite{yildizpaper}. \textit{Middle: } Model step size distribution from 32 5 second simulations. Step length is defined as the change in an MTBD position during a full step cycle. Also plotted is the head step size distribution from DeWitt \textit{et. al.} \cite{yildizpaper}. \textit{Bottom left: } Distribution of time spent in the pre-powerstroke state during a full mechanochemical cycle from the same dataset. \textit{Bottom right: } Time in the post-powerstroke state. Both bottom plots have the theoretical values for pre-powerstroke and post-powerstoke times from the DeWitt \textit{et. al.} stepping trajectory.}
\label{fig:behavior}
\end{figure}

@@

-Dynein's two feet take different steps according to yildiz/qiu
-Carter 2008 says that dynein's MTBD decides directionaolity, whereas Redwine 2012 thinks its linker motion (clarify)
-Imai 2015 has lots to say criticizing the step length studies of qiu/yildiz and have something to say about domain movement MT search
-Someone else referred to the Cianfrocco model, so maybe it's a good one to refer to (well accepted)
-It's important that our model doesn't have the powerstroke mechanism in, only observations from experiment

>>>>

%% Several computational models have been created to explain dynein's motility. \textit{Imamula et. al.} define a chemical transition model which defines all the MT- and nucleotide-bound states dynein goes through as it walks \cite{imamulamodel}. \textit{Sarlah et. al} simulate a model obeying rate constants from \textit{Imamula}, and replicate dynein's stepping trajectory \cite{sarlahmodel}. \textit{Zheng} uses normal mode analysis to simulate the dynein motor's transition from pre to post-stroke \cite{normalmodes}. To our knowledge, no computational model exists which takes into account the microscopic dynamics of protein-water interactions to test whether diffusion is enough for dynein motility. Here we show how a simple Brownian dynamics model can be used to test the mechanochemical cycle picture of dynein processivity.\\

@@

This mechanochemical cycle of chemical transitions is well supported by evidence \cite{citemore}. However, there is still relatively little known about the dynamic mechanism of domain motion dynein uses to achieve its forward motion by transitioning between these states \cite{lippert}. There exist models theorizing how dynein's domains move during its forward step, such as the winch \cite{} and powerstroke \cite{} models, which can reproduce experimental data \cite{}. However, to our knowledge, no studies exist which use protein dynamics to study if either of these dynamic mechanisms is feasible for the dynein motor to produce. Existing computational models of walking use chemical rate transitions and assume independence of steps, without simulating the precise protein dynamics \cite{}. Molecular dynamics simulations exist, but are of too short a timescale to describe dynein's forward stroke \cite{}. It is necessary to model dynein dynamically at a large timescale using physically realistic protein dynamics to fully understand how feasible models like the winch and powerstroke are. Further, such long-timescale dynamic models could be used to find what elements of dynein are sufficient to produce forward motion. Further, an accurate dynamical model would allow us to test if certain dynamical elements, such as electrostatic interactions with the microtubule \cite{longrangemt}, ring stacking \cite{sarlahmodel}, discrete microtubule binding sites \cite{trottmodel}, or elasticity of the stalk \cite{burgess-paper} are necessary for dynein's walk.

@@

\begin{figure}[tbhp]
\centering
\includegraphics[width=\linewidth]{figures/model-cartoon-simple.pdf}
\caption{\textbf{Basic geometric model in pre-stroke and post-stroke states.} \textit{Left: } Pre-stroke model of dynein with one MTBD bound to microtubule and the other free. Angles are measured from the $+\hat{x}$ direction. \textit{Right: } Post-stroke model with both MTBDs bound to microtubule. MTBD angles are relative to $+\hat{x}$ direction, and motor angles are between stalk and tail. Lengths and radii not to scale.}
\label{fig:model}
\end{figure}

\begin{figure}[tbhp]
\includegraphics[width=0.4\linewidth]{../../plots/burgess-model-figure.pdf}%
\includegraphics[width=0.6\linewidth]{../../plots/grotjahn-model-figure.pdf}%
%% \includegraphics[width=0.3\linewidth]{../../plots/crystal-model-figure.pdf}%
\caption{\textbf{Scale model compared with experimental images.} \textit{Left:} Model with angles at pre-stroke equilibrium superimposed over axonemal dynein c heavy chain micrograph in the pre-stroke ADP-Vi-bound state \cite{burgess-paper}. Interdomain lengths and radii scaled to micrograph scale bar. No microtubule is present, so $\theta_{nb}$ is arbitrary. \textit{Right: } Model with angles at pre-stroke equilibrium superimposed over density map of full-length native cytoplasmic dynein with accessory tail proteins (map ID EMD-7000) \cite{grotjahn}. Interdomain lengths and radii scaled to density map.}
\label{fig:micrographs}
\end{figure}

@@

  \begin{minipage}{\linewidth}
    \small
    \textbf{Table 1. Parameters used in simulation.} $c_b$, $c_m$ and $c_t$ are the MTBD, motor and tail spring constants, respectively. $k_b$ and $k_{ub}$ are the rates of pre-stroke to post-stroke and post-stroke to pre-stroke transitions, respectively. $L_s$ and $L_t$ are the lengths of the stalk and tail interdomain linkers. $c$ is the tension-gating factor. The $\theta$ values are the equilibrium values for each angle, where $\theta_m^{\mbox{pre}}$ is the equilibrium angle for the unbound motor in pre-stroke, whereas $\theta_m^{\mbox{post}}$ is the equilibrium angle for the bound pre-stroke motor and both post-stroke motors. $R_b$, $R_m$ and $R_t$ are the MTBD, motor and tail radii. Parameters used for all simulations unless otherwise noted.
  \end{minipage}
  \label{tab:params}

@@

\section{Model}
This dynein model is meant to capture the coarse-grained structural, dynamical and chemical properties of the full dynein complex while remaining mathematically simple. The structural features of dynein are captured through a two-dimensional geometric model of circular domains connected by rigid rods, shown in Figure \ref{fig:model}. The dynamics of dynein are captured by imposing both equilibrium and Brownian forces on each circular domain of the model. The chemical properties of dynein are captured by sporadically transitioning the model between two states: a poststroke state where both binding domains are fixed to the microtubule, and a prestroke state where one foot is free to diffuse, as has been done before \cite{zhaomodel}. The poststroke state is meant to represent dynein after rebinding free MTBD and undergoing the motor-MTBD conformational change, but before releasing ADP, according to the Tsygankov scheme \cite{tsygankovscheme}. The prestroke state represents the dynein after binding ATP, unbinding from the MT and undergoing a motor conformational change, but before rebinding the MT. ATP hydrolysis and Pi rejection are ignored by the model. These two states were chosen because ADP release is thought to be the rate limiting step for the whole cycle \cite{holzbaur1989}, and the post-unbinding conformational change is expected to happen soon after MT unbinding \cite{mogamirate}, meaning the post-unbinding conformation should dominate when unbound from MT. We constrain our model to move through only the most likely mechanochemical cycle according to \textit{Tsygankov et al} \cite{tsygankovscheme}, where a head remains in the ADP-bound post-stroke state while the other head goes through the entire mechanochemical cycle, neglecting other possible cycles. Each state feels equilibrium forces on its domains meant to mimic the structural changes which occur during the cycle \cite{burgess-paper, burgessknight}. In particular, the equilibrium angles governing a motor's tail-motor-MTBD angle change to cause a displacement of the free MTBD towards the MT minus end, producing a forward step.

@@

Values for $R$ and $L$ were acquired by aligning the dynein model with EM micrographs of the full dynein complex \cite{burgess-paper,grotjahn}, as shown in Figure \ref{fig:micrographs}. See Table \ref{table:params} for these values. $L_t$, the stem length between motor domain and tail, was calculated as the distance between the center of the EM motor domains and identified dimerization site in the tail domain \cite{}. However, studies which compare truncated GST-linked dynein lacking a tail and native dynein find the two step with similar dynamics \cite{weihongpaper-i-think}, suggesting that our model with an $L_t$ value from native dynein should be comparable with studies on truncated dynein constructs as well.

\subsection*{Equilibrium forces}
The prestroke state has four degrees of freedom: $\theta_{bb}, \theta_{bm}, \theta_{um}$ and $\theta_{ub}$. The poststroke state has two: $\theta_{nm}$ and $\theta_{fm}$. The model is free to vary across all of these degrees of freedom, depending on its state, barring steric interactions with the microtubule. These angles describe a planar model which is not capable of transverse motion. To capture the gross dynamics and interdomain oscillations of the protein, restoring forces are imposed on each angle about equilibrium angles $\theta^{eq}_{bb}, \theta^{eq}_{bm}, \theta^{eq}_{um}, \theta^{eq}_{ub}$ for prestroke and $\theta^{eq}_{nm}$ and $\theta^{eq}_{fm}$ for poststroke. Each domain $i$ has a harmonic energy $c_i\left(\theta_i-\theta^{eq}_i\right)^2$ which exerts a restoring force pushing it back to equilibrium, where $c_b$, $c_m$ and $c_t$ are spring constants for each domain. These equilibrium angles are taken to be the average angle occupied by the protein in solution, and are thus extracted from EM micrographs, shown in Figure \ref{fig:micrographs}. These angles were extracted from axonemal dynein, but structural similarity suggests they should be applicable to cytoplasmic dynein as well \cite{dynein-c-paper}. Electron microscopy studies using ADP-Vi to freeze dynein in the pre-stroke state allow for both pre- and post-stroke equilibrium angles to be found \cite{burgess-paper}.

@@

\subsection*{Binding}
Conformational changes on nucleotide binding, also known as the mechanochemical cycle \cite{cianfrocco}, are captured by transitioning the model between the prestroke and poststroke states. Binding can occur when the prestroke model's unbound foot diffuses near enough the microtubule. When within a distance of $1 A$, binding occurs at a rate of $k_b$ (Table \ref{table:params}). On binding, the prestroke model changes state to a poststroke model while retaining the positions of its domains. The previously unbound domain is transported to the microtubule.

\subsection*{Unbinding}
Unbinding occurs when a poststroke MTBD unbinds from the microtubule. This event can occur for either MTBD, and each domain has its own unbinding rate $k_{ub}$. To account for the tension-gating proposed by Yildiz \textit{et al} \cite{yildizcleary} to allow for interhead communication and bias the motor forward, we use a scheme where an MTBD is more likely to unbind based on its distance behind the other MT, which accords with experiments \cite{yildiz}. For domain $i \in \{n, f\}$ with binding domain angle $\theta_{ib}\left(\theta_{nm}, \theta_{fm}\right)$, unbinding rate is given by $k_{ub} = k^0_{ub}e^{-c\left(\theta_{ib}-\theta^{eq}_{ib}\right)}$, where $c$ is a parameter of the model. This will bias the model to unbind more preferentially for the lagging MTBD due to a higher likelihood for $\theta^{eq}_{ib} > \theta_{ib}$ for the lagging head. Values used for parameters $k^0_{ub}$ and $c$, along with $k_b$, are shown in Table \ref{table:params}.

The transitions between prestroke and poststroke events in the mechanochemical cycle \cite{cianfrocco} are captured by changing the motor equilibrium angles when a state transition occurs. The prestroke state has an unbound motor angle closer to $\pi$ than the prestroke motor\cite{burgess-paper}; thus, on unbinding, the unbound MTBD should experience a restoring force effecting the post-to-pre conformational change. On rebinding, the motors transition back to the poststroke equilibrium, producing restoring forces causing the pre-to-post conformational change. This shifting of equilibria, coupled with alternating the freedom of the active MTBD, meant to mimic the coupling of ATPase activity and MTBD affinity in the real dynein motor.

@@

  \section*{Fitting $k_b$ and $k_{ub}$}
  To fit the binding and unbinding rate constants $k^0_b$ and $k^0_{ub}$ a simple theoretical model for stepping was used, similar to one previously published \cite{myosindutyratio}. First, it is assumed that the unbinding rate of an MTBD does not depend on the binding state of the other MTBD, that is, that there is no inter-head coordination. Then the dissociation rate of the whole complex from the microtubule, $k_{dis}$, is estimated as $k_{dis} = P(ob)*k_{ub} = \frac{\bar{t}_{ob}}{\bar{t}_{step}} * \frac{1}{\bar{t}_{bb}}$, where $P(ob)$ is the probability of a dynein complex being in the prestroke state, $\bar{t}_{ob}$ and $\bar{t}_{bb}$ are the expected durations for a prestroke and poststroke state, and $\bar{t}_{step} = \bar{t}_{ob} + \bar{t}_{bb}$ is the expected step duration. Plugging in the last definition yields $k_{dis} = \frac{\bar{t}_{step} - \bar{t}_{bb}}{\bar{t}_{step}\bar{t}_{bb}}$. Introducing a new variable $\bar{t}_{run}$, the average run length of in vivo dynein, and using the relation $\bar{t}_{run} = k_{dis}^{-1}$, a new relation is found for poststroke time: $\bar{t}_{bb} = \frac{\bar{t}_{run}\bar{t}_{step}}{\bar{t}_{run}+\bar{t}_{step}}$. Using the definition of $\bar{t}_{step}$ a similar relation is found for prestroke time: $\bar{t}_{ob} = \frac{\bar{t}_{step}^2}{\bar{t}_{step}+\bar{t}_{run}}$. Finally, using the relation $\bar{t}_{step} = \frac{v}{L_{step}}$ for dynein velocity $v$ and expected tail step length $\bar{L}_{step}$, the following equations are found for prestroke and poststroke time in terms of experimental observables of velocity, run time and step length:

@@

  For estimating prestroke and poststroke times from Yildiz \textit{et al}\cite{yildiz}, a velocity of $124 nm/s$ is used. A study with a similar velocity is used to estimate $\bar{t}_{run} = 1060 nm / 134 nm / s = 7.9s$ \cite{weihongpaper}. These observables, along with $\bar{L} = 8nm$, yield estimations $\bar{t}_{ob} = 5.2*10^{-4}s$ and $\bar{t}_{bb} = 6.4 * 10^{-2}s$.

  \section*{Prestroke motion equations derivation}
  Brownian Dynamics was used to describe the motion of each domain, indexing with $i \in \{bb, bm, t, um, ub\}$, corresponding to bound-binding, bound-motor, tail, unbound-motor, unbound-binding:

@@

  \section*{Poststroke motion equations derivation}
  Brownian dynamics was also used for the poststroke case. Due to the added constraint of both binding domains being fixed to the microtubule, the model is fully described by angles $\theta_{nm}$ and $\theta_{fm}$; see Figure \ref{fig:model}. The positions of non-bound domains are defined by the following equation for $i \in \{nb, nm, t, fm, fb\}$, corresponding to near-binding, near-motor, tail, far-motor and far-binding domains:

@@

  \section*{Simulating the model}
  Simulations began in the poststroke equilibrium state. The model was time-evolved via Euler's Method using analytic expressions for $\{\dot{\theta}_{bb}, \dot{\theta}_{bm}, \dot{\theta}_{um}, \dot{\theta}_{ub}\}$, using a timestep of $dt = 10^{-10}s$. In the prestroke state, during each timestep the model had a $k_b*dt$ probability of transitioning into the poststroke state, where $k_b = k_b^0\left(1-H\left(X_{uby}-0.1\right)\right)$, where H is the Heaviside function. On transition to poststroke, proper $\theta_{nm}$ and $\theta_{fm}$ angles and proper interhead separation $L$ were calculated to yield a poststroke model with identical cartesian positions to the old poststroke, except with the previously unbound binding domain moved to $y=0$. Once in poststroke, the model was time-evolved via Euler's Method using analytic expressions for $\{\dot{\theta}_{nm}$ and $\dot{\theta}_{fm}\}$ and a timestep of $dt=10^{-10}s$. Each timestep, the poststroke model had a $k_{ub}*dt$ chance of transitioning to prestroke, where $k_{ub} = k^0_{ub}e^{-c(\theta-\theta_{eq})}$. On transitioning, proper values of $\{\theta_{bb}, \theta_{bm}, \theta_{um}, \theta_{ub}\}$ were calculated to create a prestroke model with identical cartesian positions to the old poststroke model.

@@

Motion equations were derived for the pre-stroke and post-stroke states of the model. The model was simulated by starting in the equilibrium post-stroke position and using Euler's method to time-evolve the model, using a timestep of 1e-10 \textit{s}. The model was transitioned between post-stroke and pre-stroke states using unbinding and binding rates. Model parameters were estimated from various experimental results. Spring constants were chosen to fit the model stepping size to single-molecule head-labelled experiments\cite{yildizpaper}, as outlined in Figure S\ref{fig:supp-model-fit}. Model rate constants were fit to motor velocities from the same experiment, along with theoretical calculations for the prestroke and poststroke times outlined in the Methods section. The tension-gating parameter $c$ was fit to the leading/lagging unbinding likelihood versus interhead separation from the same experiment, as shown in Figure \ref{fig:tensiongating}. Thirty-two 5-second simulations were run, each initialized with a different random seed. Each simulation was forward directed, took both forward and backward steps, and displayed stepping dynamics similar to that seen in experiment.

@@

The model was then compared to TIRF studies of dynein stepping in DeWitt \textit{et. al.} \cite{yildizpaper}. Figure \ref{fig:behavior}.a shows 8 representative model trajectories compared to experimental dynein. The eight simulations track experimental very well over 5 seconds, and average to the same position. The eight simulations end up distributed over a 200 \textit{nm} area, showing the variation in the model stepping extends to a variation in velocity in the long term. The step size distribution for the 32 simulations are shown in Figure \ref{fig:behavior}.b). The model is capable of taking both positive and negative steps, with steps extending from -40 \textit{nm} to 50 \textit{nm}, matching experimental dynein. The most probable step is 12 \textit{nm}, with an average step size of 6.2 \textit{nm}. This matches the peak experimental step size window of 7-13 \textit{nm}. The pre-stroke and post-stroke state occupation times are shown in Figure \ref{fig:behavior}.c and d.). The model prefers long pre-stroke times, but is capable of taking very fast steps. The post-stroke time has a heavy-tailed distribution. Experimental technique does not resolve time spent in pre- and post-stroke, but we outline a technique for calculating these values for experimental trajectories in Methods. These theoretical values correspond well to model behavior, and square with experimental findings that dynein is almost always in the post-stroke state \cite{imaiburgess}.

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\linewidth]{../../plots/paper_model_behavior}
\caption{\textbf{Stepping dynamics of model.} \textit{Top: } Average position of near and far MTBDs over 5 seconds of simulation and 8 random seeds. Trajectory compared with single-motor-labelled cytoplasmic dynein from DeWitt \textit{et. al.} \cite{yildizpaper}. \textit{Middle: } Model step size distribution from 32 5 second simulations. Step length is defined as the change in an MTBD position during a full step cycle. Also plotted is the head step size distribution from DeWitt \textit{et. al.} \cite{yildizpaper}. \textit{Bottom left: } Distribution of time spent in the pre-stroke state during a full mechanochemical cycle from the same dataset. \textit{Bottom right: } Time in the post-stroke state. Both bottom plots have the theoretical values for pre-stroke and post-powerstoke times from the DeWitt \textit{et. al.} stepping trajectory.}
\label{fig:behavior}
\end{figure}

@@
